HO unification from object language to meta language

### 2024-03-30 paper/main.tex
free mathcal P for subterm

\newcommand{\printEntry}[1]{
  \listlength{#1}%
  \ifthenelse{\equal{0}{\thellength}}{}{
    \item[Entry pb:] 
    \foreach \x [count=\i] in #1
      {%
        \begin{minipage}{8cm}%
          $\mathcal{P}_\i : \x$ %
        \end{minipage}%
        \ifthenelse{\equal{\i}{\thellength}}{}{\\}%
      }%
  }%
}
\newcommand{\printItemComp}[3]{
  \listlength{#3}%
  \ifthenelse{\equal{0}{\thellength}}{}{
    \item[#1] \printMinipage{#2}{#3}
  }%
}

\newcommand{\printAll}[4]{
  \noindent%
  \begin{description}[leftmargin=!,labelwidth=\widthof{\bfseries Mapping:}]
    \printEntry{#1}%
    \printItemComp{Compil:}{\mathcal{T}}{#2}%
    \printItemComp{Mapping:}{m}{#3}%
    \printItemComp{Links:}{l}{#4}
  \end{description}
}

@@

We represent a logic program \emph{run} in \Fo{} as
a list \emph{steps} $p$ of length $\mathcal{N}$. Each step is a
unification problem between terms $\mathcal{S}_{p_l}$ and
$\mathcal{S}_{p_r}$ taken from the set of all terms $\mathcal{S}$.
The composition of these steps starting from the
empty substitution $\rho_0$ produces the final
substitution $\rho_\mathcal{N}$.
\footnote{If the same rule is used multiple time in a run we
just consider as many copies as needed of the terms composing the
rules, with fresh unification variables each time}
The initial here $\rho_0$ is the empty substitution
%
\newcommand{\progress}{\ensuremath{\mathrm{progress}}\xspace}
\newcommand{\fstep}{\ensuremath{\mathrm{fstep}}\xspace}
\newcommand{\hstep}{\ensuremath{\mathrm{hstep}}\xspace}
\newcommand{\frun}{\ensuremath{\mathrm{frun}}\xspace}
\newcommand{\hrun}{\ensuremath{\mathrm{hrun}}\xspace}
\newcommand{\stepF}[4]{\ensuremath{\fstep(#1,#2,#3) \mapsto #4}}
\newcommand{\stepFD}[5]{%
\ensuremath{#3 #1_{#2_l} \Uo #3 #1_{#2_r} \mapsto #4 \land #5 = #3 \cup #4}}
\newcommand{\stepH}[6]{\ensuremath{\hstep(#1,#2,#3,#4) \mapsto (#5, #6)}}
\newcommand{\stepHD}[6]{\ensuremath{%
#3 #1_{#2_l} \Ue #3 #1_{#2_r} \mapsto #4 \land \progress(#6,#3 \cup #4) \mapsto (#6',#5)}}
\newcommand{\runF}[3]{\ensuremath{\frun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runFD}[2]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepF{#1}{p}{\rho_{p-1}}{\rho_{p}}}}
\newcommand{\runH}[3]{\ensuremath{\hrun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runHD}[3]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepH{#1}{p}{\sigma_{p-1}}{#3_{p-1}}{\sigma_{p}}{#3_p}}}
\newcommand{\deff}{\ensuremath{\stackrel{{\scriptscriptstyle def}}{=\!=\!\!\!=}}}
\newcommand{\linkStore}{\ensuremath{\mathbb{L}}\xspace}
\newcommand{\mapStore}{\ensuremath{\mathbb{M}}\xspace}
%
$$
\begin{array}{l}
\stepF{\mathcal{S}}{p}{\rho}{\rho''}
\deff
\stepFD{\mathcal{S}}{p}{\rho}{\rho'}{\rho''}\vspace{2pt}\\
\runF{\mathcal{S}}{\mathcal{N}}{\rho}
\deff
\runFD{\mathcal{S}}{\mathcal{N}}
\end{array}
$$

We simulate each run in \Fo{} with a run in \Ho as follows.
Note that $\sigma_0$ is the empty substitution.
$$
\begin{array}{l}
\stepH{\mathcal{T}}{p}{\sigma}{\linkStore}{\sigma''}{\linkStore'} \deff\vspace{2pt}\\
  \qquad\stepHD{\mathcal{T}}{p}{\sigma}{\sigma'}{\sigma''}{\linkStore}\vspace{2pt}\\
\runH{\mathcal{S}}{\mathcal{N}}{\rho} \deff \vspace{2pt}\\
  \qquad \mathcal{T} \times \mapStore \times \linkStore_0 = \{ (t_j,m_j,l_j) | s_j \in \mathcal{S}, \C{s_j}{t_j}{m_j}{l_j} \}\vspace{2pt}\\
  \qquad \runHD{\mathcal{T}}{\mathcal{N}}{\linkStore}\vspace{2pt}\\
  \qquad \D{\sigma_{\mathcal{N}}}{\mapStore}{\linkStore_{\mathcal{N}}}{\rho_{\mathcal{N}}}
\end{array}
$$

@@

\begin{proposition}[Simulation]\label{prop:simulation}
$\forall \mathcal{S}, \forall \mathcal{N},$
$$
  \runF{\mathcal{S}}{\mathcal{N}}{\rho}
  \Leftrightarrow
  \runH{\mathcal{S}}{\mathcal{N}}{\rho}
$$
\end{proposition}

@@

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of~ \hrun, if $~\mathcal{T} \subseteq \llambda$ we have that
$\forall p \in 1 \ldots \mathcal{N},$
$$
\stepF{\mathcal{S}}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\mathcal{T}}{p}{\sigma_{p-1}}{\linkStore}{\sigma_p}{\_}
$$
\end{proposition}
\noindent
In particular this property guarantees that a \emph{failure} in the \Fo{} run
is matched by a failure in \Ho{} \emph{at the same step}. We consider this
property very important from a practical point of view since it guarantees
that the execution traces are strongly related and in turn this enables a user
to debug a logic program in \Fo{} by looking at its execution trace in
\Ho{}.

@@

We can define $s_1 \Uo{} s_2$ by specializing the code of \hrun{} to
$\mathcal{S} = \{ s_1, s_2 \}$ as follows:
%
$$
\begin{array}{l}
s_1 \Uo s_2 \mapsto \rho \deff \vspace{2pt}\\
\quad\C{s_1}{t_1}{m_1}{l_1} \land \C{s_2}{t_2}{m_2}{l_2}\vspace{2pt}\\
\quad    t_1 \Ue t_2 \mapsto \sigma' \land
    \progress(\{l_1,l_2\},\sigma') \mapsto (L,\sigma'') \land\vspace{2pt}\\
\quad \D{\sigma''}{\{m_1,m_2\}}{L}{\rho}
\end{array}
$$

@@

\newcommand{\subterm}[1]{\ensuremath{\mathcal{P}(#1)}}
\begin{definition}[Subterms \subterm{t}] The set of sub terms of $t$ is the
  largest set $\mathcal{P(t)}$ that can be obtained by the following rules.
$$
\begin{array}{l}
t \in \subterm{t}\\
t = f t_1\ldots t_n \Rightarrow \subterm{t_i} \subseteq \subterm{t} \land f \in \subterm{t}\\
t = \lambda x.t' \Rightarrow \subterm{t'} \subseteq \subterm{t}\\
\end{array}
$$
\end{definition}

@@

\begin{proposition}[\wellb{}-preservation]\label{prop:nf}
$\forall \mathcal{T}, \forall \linkStore, \forall p, \forall \sigma, \forall \sigma'$
$$
\begin{array}{l}
\wellb(\sigma\mathcal{T}) \land
  \sigma\mathcal{T}_{p_l} \Ue \sigma\mathcal{T}_{p_r} \mapsto {\sigma'}
  \Rightarrow \wellb(\sigma' \mathcal{T})\\
\wellb(\sigma\mathcal{T}) \land
  \progress(\linkStore,\sigma) \mapsto (\_,\sigma')
  \Rightarrow \wellb(\sigma' \mathcal{T})
\end{array}
$$
\end{proposition}

@@

\section{Handling of \maybeeta}\label{sec:eta}
\marginpar{TODO: fix syntax}
$\eta$-reduction is an equivalence relation where a term of the form
$\lambda x.t \appsep x$ can be converted to $t$ any time $x$ does not occur as a
free variable in $t$. We call $\lambda x. t \appsep x$ the $\eta$-expansion of $t$. 
\printAll
  {{\elpiIn{flam x\ fapp [fuva X, x]} \Uo \elpiIn{fcon"f"}}}
  {{\lambda x. X'_x \Ue f}}
  {{\mapping{\elpiIn{X}}{X'}{1}}}
  {{}}
Following the implementation of \elpiIn{comp} given in
\cref{sec:compilation}, the unification problem $\mathcal{P}_1$ given 
above, is compiled into the unification problem $\mathcal{T}_1$. 
We can note that all the terms appearing in $\mathcal{P}_1$ are
unifiable by \Uo, but, despite this, \Ue, is not able to solve $\mathcal{T}_1$:
its left and right hand sides have both different rigid head:
$\lambda x.A'_x$ corresponds to the \Ho term \elpiIn{lam x\ app[con "f", x]} 
and $f$ corresponds to \elpiIn{con"f"}. 
This failure is motivated by the fact that \elpiIn{flam x\ fapp [fuva A, x]}
is a term belonging to \maybeeta.

@@

Once a \linketa is activated, it can be removed from \linkStore,
otherwise, the link is kept for a further iteration of \elpiIn{progress}. Note
that this link progression enforce
\cref{prop:nf,inv:linklhs,inv:link-eta-right}: we never commit a term in
the \Ho substitution, since we make unification only when we know that the 
terms are no more \maybeeta, and when \lhs is no more a variable or \rhs is 
no more a \maybeeta, the link is removed from \linkStore.
% \printAll
%   {{\elpiIn{flam x\ fapp [fuva X, x] ~\Uo~fcon"f"}}}
%   {{A \Ue f}}
%   {{\mapping{X}{X'}{1}}}
%   {{\linketaM{}{A}{\lambda x.X'_{x}}}}
$$
\begin{array}{l}
  \mathcal{P} = \{~ \lambda x.X \appsep x \Uo f ~\}\\
  \mathcal{T} = \{~ A \Ue f ~\}\\
  \mathbb{M} = \{~ \mapping{X}{B}{1} ~\}\\
  \linkStore \;= \{~ \linketaM{}{A}{\lambda x.B_{x ~\}}}
\end{array}
$$

The example above shows the new compilation of the unification problem given at
the beginning of \cref{sec:eta}. This time, we see that the the left hand side
$t$ of $\mathcal{P}_1$ has been detected as a \maybeeta and replaced with the
fresh variable $A$. Moreover, \linkStore contains the link $l_1$ connecting $A$
with $\lambda x.X'_{x}$, the compiled version of $t$. After the resolution of
$\mathcal{T}_1$, $A$ is assigned to $f$. Therefore \lhs of $l_1$ is a term with
\elpiIn{con} as constructor. This means that \rhs of $l_1$ is a $\eta$-expansion
and therefore we can unify $\lambda x.f x$ with \rhs, which instantiate $X'_x$
to $f \appsep x$. $l_1$ is removed from \linkStore which is now empty,
\elpiIn{progress} terminate, and the decompilation, will instantiate the $X$
variable to $f$ (which is the $\eta$-contracted version of $\lambda x.f \appsep
x$).

@@

\printAll
  {{\elpiIn{flam x\ flam y\ fapp [fuva X, y, x] == flam x\ flam y\ x},
    \elpiIn{flam x\ fapp [fcon"f", fapp [fuva X, x], x] == fuva Y}}}
  {{E0 == \lambda x.\lambda y.x,
    E3 == E5}}
  {{\mapping{Y}{E5}{0},
    \mapping{X}{E2}{2}}}
  {{\linketaM{c0}{E4_{c0}}{\lambda x.E2_{c0 x}},
    \linketaM{}{E3}{\lambda x.(f~E4_{x}~x)},
    \linketaM{}{E0}{\lambda x.E1_{x}},
    \linketaM{c0}{E1_{c0}}{\lambda x.E2_{x c0}}}}
$$
{\arraycolsep=2pt \begin{array}{rrclrcl}
    \mathcal{P} = \{ &
       \lambda x.\lambda y.X \appsep y \appsep x & \Uo & \lambda x.\lambda y. x & \quad
       \lambda x. f \appsep (X \appsep x) \appsep x & \Uo&  Y ~\}\\
    \mathcal{T} = \{ &
       A &\Ue& \lambda x.\lambda y.x & \quad
       D &\Ue& F ~\}\\
    \mathbb{M} = \{ & \multicolumn{6}{l}{
      \mapping{Y}{F}{0} \quad
      \mapping{X}{C}{2} ~\}}\\
    \linkStore \;= \Big\{  &
      \multicolumn{6}{l}{ \begin{array}{ll}
          \linketaM{}{A}{\lambda x.B_{x}} & \quad
          \linketaM{x}{B_{x}}{\lambda y.C_{y x}}\\
          \linketaM{}{D}{\lambda x.(f~E_{x}~x)} & \quad
          \linketaM{y}{E_{y}}{\lambda w.C_{y w}}
      \end{array} ~\Big\}}
\end{array}}
$$
\marginpar{aggiustare i nomi delle variabili nel testo}
We see that the \elpiIn{maybe-eta} as detected $\lambda xy.F \appsep y \appsep
x$ and $\lambda x.f\appsep (F \appsep x) \appsep x$ and replaced them with
respectively the \Ho vars $X$ and $Z$. $X$ is linked with $\lambda x.Y_x$, $Y$
has arity $1$ and is $\eta$-linked with $\lambda y.H\appsep y \appsep x$ and $Z$
is linked to the term $\lambda x.f \appsep G_{x}~x$. However, the mapping
returned by the compilation, does not breaks \cref{inv:map-bijective}: the \Fo{}
variable $F$ is mapped to two different \Ho variables. To address this problem
and enforce \cref{inv:map-bijective}, we clean the mapping with a second phase
after the compilation. This phase is called \elpiIn{map-deduplication}.

>>>>

\newcommand{\printEntry}[1]{
  \listlength{#1}%
  \ifthenelse{\equal{0}{\thellength}}{}{
    \item[Entry pb:] 
    \foreach \x [count=\i] in #1
      {%
        \begin{minipage}{8cm}%
          $\mathbb{P}_\i : \x$ %
        \end{minipage}%
        \ifthenelse{\equal{\i}{\thellength}}{}{\\}%
      }%
  }%
}
\newcommand{\printItemComp}[3]{
  \listlength{#3}%
  \ifthenelse{\equal{0}{\thellength}}{}{
    \item[#1] \printMinipage{#2}{#3}
  }%
}

\newcommand{\printAll}[4]{
  \noindent%
  \begin{description}[leftmargin=!,labelwidth=\widthof{\bfseries Mapping:}]
    \printEntry{#1}%
    \printItemComp{Compil:}{\mathbb{T}}{#2}%
    \printItemComp{Mapping:}{m}{#3}%
    \printItemComp{Links:}{l}{#4}
  \end{description}
}

@@

We represent a logic program \emph{run} in \Fo{} as
a list \emph{steps} $p$ of length $\mathcal{N}$. Each step is a
unification problem between terms $\mathbb{S}_{p_l}$ and
$\mathbb{S}_{p_r}$ taken from the set of all terms $\mathbb{S}$.
The composition of these steps starting from the
empty substitution $\rho_0$ produces the final
substitution $\rho_\mathcal{N}$.
\footnote{If the same rule is used multiple time in a run we
just consider as many copies as needed of the terms composing the
rules, with fresh unification variables each time}
The initial here $\rho_0$ is the empty substitution
%
\newcommand{\progress}{\ensuremath{\mathrm{progress}}\xspace}
\newcommand{\fstep}{\ensuremath{\mathrm{fstep}}\xspace}
\newcommand{\hstep}{\ensuremath{\mathrm{hstep}}\xspace}
\newcommand{\frun}{\ensuremath{\mathrm{frun}}\xspace}
\newcommand{\hrun}{\ensuremath{\mathrm{hrun}}\xspace}
\newcommand{\stepF}[4]{\ensuremath{\fstep(#1,#2,#3) \mapsto #4}}
\newcommand{\stepFD}[5]{%
\ensuremath{#3 #1_{#2_l} \Uo #3 #1_{#2_r} \mapsto #4 \land #5 = #3 \cup #4}}
\newcommand{\stepH}[6]{\ensuremath{\hstep(#1,#2,#3,#4) \mapsto (#5, #6)}}
\newcommand{\stepHD}[6]{\ensuremath{%
#3 #1_{#2_l} \Ue #3 #1_{#2_r} \mapsto #4 \land \progress(#6,#3 \cup #4) \mapsto (#6',#5)}}
\newcommand{\runF}[3]{\ensuremath{\frun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runFD}[2]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepF{#1}{p}{\rho_{p-1}}{\rho_{p}}}}
\newcommand{\runH}[3]{\ensuremath{\hrun(#1,#2) \mapsto #3_{#2}}}
\newcommand{\runHD}[3]{\ensuremath{%
\bigwedge_{p = 1}^{#2} \stepH{#1}{p}{\sigma_{p-1}}{#3_{p-1}}{\sigma_{p}}{#3_p}}}
\newcommand{\deff}{\ensuremath{\stackrel{{\scriptscriptstyle def}}{=\!=\!\!\!=}}}
\newcommand{\linkStore}{\ensuremath{\mathbb{L}}\xspace}
\newcommand{\mapStore}{\ensuremath{\mathbb{M}}\xspace}
%
$$
\begin{array}{l}
\stepF{\mathbb{S}}{p}{\rho}{\rho''}
\deff
\stepFD{\mathbb{S}}{p}{\rho}{\rho'}{\rho''}\vspace{2pt}\\
\runF{\mathbb{S}}{\mathcal{N}}{\rho}
\deff
\runFD{\mathbb{S}}{\mathcal{N}}
\end{array}
$$

We simulate each run in \Fo{} with a run in \Ho as follows.
Note that $\sigma_0$ is the empty substitution.
$$
\begin{array}{l}
\stepH{\mathbb{T}}{p}{\sigma}{\linkStore}{\sigma''}{\linkStore'} \deff\vspace{2pt}\\
  \qquad\stepHD{\mathbb{T}}{p}{\sigma}{\sigma'}{\sigma''}{\linkStore}\vspace{2pt}\\
\runH{\mathbb{S}}{\mathcal{N}}{\rho} \deff \vspace{2pt}\\
  \qquad \mathbb{T} \times \mapStore \times \linkStore_0 = \{ (t_j,m_j,l_j) | s_j \in \mathbb{S}, \C{s_j}{t_j}{m_j}{l_j} \}\vspace{2pt}\\
  \qquad \runHD{\mathbb{T}}{\mathcal{N}}{\linkStore}\vspace{2pt}\\
  \qquad \D{\sigma_{\mathcal{N}}}{\mapStore}{\linkStore_{\mathcal{N}}}{\rho_{\mathcal{N}}}
\end{array}
$$

@@

\begin{proposition}[Simulation]\label{prop:simulation}
$\forall \mathbb{S}, \forall \mathcal{N},$
$$
  \runF{\mathbb{S}}{\mathcal{N}}{\rho}
  \Leftrightarrow
  \runH{\mathbb{S}}{\mathcal{N}}{\rho}
$$
\end{proposition}

@@

\begin{proposition}[Simulation fidelity]\label{prop:fidelity}
In the context of~ \hrun, if $~\mathbb{T} \subseteq \llambda$ we have that
$\forall p \in 1 \ldots \mathcal{N},$
$$
\stepF{\mathbb{S}}{p}{\rho_{p-1}}{\rho_p}
\Leftrightarrow
\stepH{\mathbb{T}}{p}{\sigma_{p-1}}{\linkStore}{\sigma_p}{\_}
$$
\end{proposition}
\noindent
In particular this property guarantees that a \emph{failure} in the \Fo{} run
is matched by a failure in \Ho{} \emph{at the same step}. We consider this
property very important from a practical point of view since it guarantees
that the execution traces are strongly related and in turn this enables a user
to debug a logic program in \Fo{} by looking at its execution trace in
\Ho{}.

@@

We can define $s_1 \Uo{} s_2$ by specializing the code of \hrun{} to
$\mathbb{S} = \{ s_1, s_2 \}$ as follows:
%
$$
\begin{array}{l}
s_1 \Uo s_2 \mapsto \rho \deff \vspace{2pt}\\
\quad\C{s_1}{t_1}{m_1}{l_1} \land \C{s_2}{t_2}{m_2}{l_2}\vspace{2pt}\\
\quad    t_1 \Ue t_2 \mapsto \sigma' \land
    \progress(\{l_1,l_2\},\sigma') \mapsto (L,\sigma'') \land\vspace{2pt}\\
\quad \D{\sigma''}{\{m_1,m_2\}}{L}{\rho}
\end{array}
$$

@@

\newcommand{\subterm}[1]{\ensuremath{\mathcal{P}(#1)}}
\begin{definition}[Subterms \subterm{t}] The set of sub terms of $t$ is the
  largest set $\\subterm{t}$ that can be obtained by the following rules.
$$
\begin{array}{l}
t \in \subterm{t}\\
t = f t_1\ldots t_n \Rightarrow \subterm{t_i} \subseteq \subterm{t} \land f \in \subterm{t}\\
t = \lambda x.t' \Rightarrow \subterm{t'} \subseteq \subterm{t}\\
\end{array}
$$
\end{definition}

@@

\begin{proposition}[\wellb{}-preservation]\label{prop:nf}
$\forall \mathbb{T}, \forall \linkStore, \forall p, \forall \sigma, \forall \sigma'$
$$
\begin{array}{l}
\wellb(\sigma\mathbb{T}) \land
  \sigma\mathbb{T}_{p_l} \Ue \sigma\mathbb{T}_{p_r} \mapsto {\sigma'}
  \Rightarrow \wellb(\sigma' \mathbb{T})\\
\wellb(\sigma\mathbb{T}) \land
  \progress(\linkStore,\sigma) \mapsto (\_,\sigma')
  \Rightarrow \wellb(\sigma' \mathbb{T})
\end{array}
$$
\end{proposition}

@@

\section{Handling of \maybeeta}\label{sec:eta}
\marginpar{TODO: fix syntax}
$\eta$-reduction is an equivalence relation where a term of the form
$\lambda x.t \appsep x$ can be converted to $t$ any time $x$ does not occur as a
free variable in $t$. We call $\lambda x. t \appsep x$ the $\eta$-expansion of $t$. 
\printAll
  {{\elpiIn{flam x\ fapp [fuva X, x]} \Uo \elpiIn{fcon"f"}}}
  {{\lambda x. X'_x \Ue f}}
  {{\mapping{\elpiIn{X}}{X'}{1}}}
  {{}}
Following the implementation of \elpiIn{comp} given in
\cref{sec:compilation}, the unification problem $\mathbb{P}_1$ given 
above, is compiled into the unification problem $\mathbb{T}_1$. 
We can note that all the terms appearing in $\mathbb{P}_1$ are
unifiable by \Uo, but, despite this, \Ue, is not able to solve $\mathbb{T}_1$:
its left and right hand sides have both different rigid head:
$\lambda x.A'_x$ corresponds to the \Ho term \elpiIn{lam x\ app[con "f", x]} 
and $f$ corresponds to \elpiIn{con"f"}. 
This failure is motivated by the fact that \elpiIn{flam x\ fapp [fuva A, x]}
is a term belonging to \maybeeta.

@@

Once a \linketa is activated, it can be removed from \linkStore,
otherwise, the link is kept for a further iteration of \elpiIn{progress}. Note
that this link progression enforce
\cref{prop:nf,inv:linklhs,inv:link-eta-right}: we never commit a term in
the \Ho substitution, since we make unification only when we know that the 
terms are no more \maybeeta, and when \lhs is no more a variable or \rhs is 
no more a \maybeeta, the link is removed from \linkStore.
% \printAll
%   {{\elpiIn{flam x\ fapp [fuva X, x] ~\Uo~fcon"f"}}}
%   {{A \Ue f}}
%   {{\mapping{X}{X'}{1}}}
%   {{\linketaM{}{A}{\lambda x.X'_{x}}}}
$$
\begin{array}{l}
  \mathbb{P} = \{~ \lambda x.X \appsep x \Uo f ~\}\\
  \mathbb{T} = \{~ A \Ue f ~\}\\
  \mathbb{M} = \{~ \mapping{X}{B}{1} ~\}\\
  \linkStore \;= \{~ \linketaM{}{A}{\lambda x.B_{x ~\}}}
\end{array}
$$

The example above shows the new compilation of the unification problem given at
the beginning of \cref{sec:eta}. This time, we see that the the left hand side
$t$ of $\mathbb{P}_1$ has been detected as a \maybeeta and replaced with the
fresh variable $A$. Moreover, \linkStore contains the link $l_1$ connecting $A$
with $\lambda x.X'_{x}$, the compiled version of $t$. After the resolution of
$\mathbb{T}_1$, $A$ is assigned to $f$. Therefore \lhs of $l_1$ is a term with
\elpiIn{con} as constructor. This means that \rhs of $l_1$ is a $\eta$-expansion
and therefore we can unify $\lambda x.f x$ with \rhs, which instantiate $X'_x$
to $f \appsep x$. $l_1$ is removed from \linkStore which is now empty,
\elpiIn{progress} terminate, and the decompilation, will instantiate the $X$
variable to $f$ (which is the $\eta$-contracted version of $\lambda x.f \appsep
x$).

@@

\printAll
  {{\elpiIn{flam x\ flam y\ fapp [fuva X, y, x] == flam x\ flam y\ x},
    \elpiIn{flam x\ fapp [fcon"f", fapp [fuva X, x], x] == fuva Y}}}
  {{E0 == \lambda x.\lambda y.x,
    E3 == E5}}
  {{\mapping{Y}{E5}{0},
    \mapping{X}{E2}{2}}}
  {{\linketaM{c0}{E4_{c0}}{\lambda x.E2_{c0 x}},
    \linketaM{}{E3}{\lambda x.(f~E4_{x}~x)},
    \linketaM{}{E0}{\lambda x.E1_{x}},
    \linketaM{c0}{E1_{c0}}{\lambda x.E2_{x c0}}}}
$$
{\arraycolsep=2pt \begin{array}{rrclrcl}
    \mathbb{P} = \{ &
       \lambda x.\lambda y.X \appsep y \appsep x & \Uo & \lambda x.\lambda y. x & \quad
       \lambda x. f \appsep (X \appsep x) \appsep x & \Uo&  Y ~\}\\
    \mathbb{T} = \{ &
       A &\Ue& \lambda x.\lambda y.x & \quad
       D &\Ue& F ~\}\\
    \mathbb{M} = \{ & \multicolumn{6}{l}{
      \mapping{Y}{F}{0} \quad
      \mapping{X}{C}{2} ~\}}\\
    \linkStore \;= \Big\{  &
      \multicolumn{6}{l}{ \begin{array}{ll}
          \linketaM{}{A}{\lambda x.B_{x}} & \quad
          \linketaM{x}{B_{x}}{\lambda y.C_{y x}}\\
          \linketaM{}{D}{\lambda x.(f~E_{x}~x)} & \quad
          \linketaM{y}{E_{y}}{\lambda w.C_{y w}}
      \end{array} ~\Big\}}
\end{array}}
$$
\marginpar{aggiustare i nomi delle variabili nel testo}
We see that the \elpiIn{maybe-eta} as detected $\lambda xy.F \appsep y \appsep
x$ and $\lambda x.f\appsep (F \appsep x) \appsep x$ and replaced them with
respectively the \Ho vars $X$ and $Z$. $X$ is linked with $\lambda x.Y_x$, $Y$
has arity $1$ and is $\eta$-linked with $\lambda y.H\appsep y \appsep x$ and $Z$
is linked to the term $\lambda x.f \appsep G_{x}~x$. However, the mapping
returned by the compilation, does not breaks \cref{inv:map-bijective}: the \Fo{}
variable $F$ is mapped to two different \Ho variables. To address this problem
and enforce \cref{inv:map-bijective}, we clean the mapping with a second phase
after the compilation. This phase is called \elpiIn{map-deduplication}.
